---
output: html_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%",
  out.height = "80%"
)

library(tidyverse)
library(ggIFS)
```

# ggIFS

<!-- badges: start -->

<!-- badges: end -->

ggplot2 is the standard package for creating graphs in R. ggIFS is a package that adds on to ggplot2 to make graphs that are consistent with the IFS's branding.

For those that are not familiar with the structure of ggplot2, 'theme' and colours are dealt with separately. Theme refers on the graph that is not the data - i.e. the axis ticks and lines, the plot background, the title, the size of the text, etc. Adding `theme_ifs()` to a ggplot2 graph adds the standard IFS theme elements.

Colours are applied separately. Unhelpfully, ggplot2 has two ways to apply colour: 'fill' and 'colour'. Fill is for things you want to fill, such as bar graphs and area graphs, and colour is for things you want to outline, such as line graphs and scatter graphs. Colours are applied separately. `scale_colour_ifs()` and `scale_fill_ifs()` are wrappers for `scale_colour_manual()` and `scale_fill_manual()` respectively, and apply the IFS colour scheme to graphs.

Click [here](https://ggplot2-book.org/) for a more general guide to ggplot2.

## Installation

You can install the development version of ggIFS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vedanthnair/ggIFS")
```

## Examples

### Standard bar chart

This is an example which applies the IFS theme and colour to a bar plot.

```{r example}
# Creating an example dataframe, which has the mean variable for each species. Iris is a built-in dataset used for examples. 
iris_clean <- iris %>%
  group_by(Species) %>%
  summarize(across(.cols = everything(), mean, na.rm = T)) %>% 
  pivot_longer(cols = !Species) %>%
  arrange(name)

head(iris_clean)
```

Without putting in any arguments, `scale_fill_ifs()` (we are using fill because we want a bar chart), will automatically apply the IFS standard colour palette (mid-primary colours).

```{r example2, fig.height= 3, fig.width = 5}

ggplot(iris2) +
  geom_bar(aes(x = Species, y = value, fill = name), stat = "identity") +
  # theme_ifs() adds the theme elements
  theme_ifs() +
  # scale_fill_ifs() without any options adds the colours
  scale_fill_ifs()
```

### Standard line graph

Creating a line graph is pretty similar to creating a bar chart, except you have to use the colour aesthetic and `scale_colour_ifs()` .

```{r , fig.height= 3, fig.width = 5}
#create another dummy dataset
iris_clean2 <- iris %>%
  mutate(row_number = row_number()) %>%
  pivot_longer(cols = Sepal.Length:Petal.Width)

ggplot(iris_clean2) +
  # Note - row thickness is not dealt with by theme, so has to be set manually
  geom_line(aes(x = row_number, y = value, colour = name), size = 0.75) +
  theme_ifs() +
  scale_colour_ifs() 
```

### Choosing a specific palette

We can choose a custom palette, e.g. greens. Palettes can be greens, greys, yellows, reds, blues, purples.

```{r example3, fig.height= 3, fig.width = 5}
ggplot(iris2) +
  geom_bar(aes(x = Species, y = value, fill = name), stat = "identity") +
  theme_ifs() +
  scale_fill_ifs(palette = "greens")
```

### User-specified palette

We can choose custom colour values, using the 'values' argument. The potential values are dark_X\_2, dark_X\_1, mid_X, light_X\_1, light_X\_2, light_X\_3, substituting X for green, grey, yellow, red, blue or purple (following the IFS naming convention).

```{r example4, fig.height= 3, fig.width = 5}
ggplot(iris2) +
  geom_bar(aes(x = Species, y = value, fill = name), stat = "identity") +
  theme_ifs() +
  scale_fill_ifs(values = c("Petal.Length" = "dark_green_2", "Petal.Width" = "dark_purple_1", "Sepal.Length" = "mid_yellow", "Sepal.Width" = "light_grey_3"))

```

### Bar Chart - Single Colours

The `scale_fill_ifs` and `scale_colour_ifs` commands are designed to map to the `fill` or `colour` aesthetics. This makes it awkward if you want a graph with a single colour, as there will be nothing to put in the `fill/colour` aesthetic.

There are two options to get around this. The first is just to set `fill or colour = ""` in the `aes()` call, and then manually suppress the legend ggplot will try to create.

```{r , fig.height= 3, fig.width = 5}
ggplot(iris2) +
  geom_bar(aes(x = Species, y = value, fill = ""), stat = "identity") +
  theme_ifs() +
  theme(legend.position = "none") +
  scale_fill_ifs()

```

Alternatively, you can forget about using `scale_fill_ifs()` and just manually specify the colour. The hex value for the standard IFS is #309E75. Because we are manually specifying the colour and not linking it to a variable, ggplot needs you to put it outside of the `aes()` call.

```{r eval=FALSE}
# This will create the same code as above. 
ggplot(iris2) +
  geom_bar(aes(x = Species, y = value), fill = "#307E75", stat = "identity")+
  theme_ifs()
```

### Scatter Plot

You can also add gradients, in which case the palette or user-specified values are taken as start, end and mid points of the gradient.

```{r example6, fig.height= 3, fig.width = 5}
ggplot(iris) +
  geom_point(aes(x = Sepal.Length, y = Sepal.Width, col = Petal.Width)) +
  theme_ifs() +
  scale_colour_ifs(palette = "greens", gradient = TRUE)

```

## Saving

Unlike Excel, R graph elements do not scale with size. Say you originally save a graph as 50mm x 100mm. Later, you decide to save the same graph as 100mm x 200mm. Although the graph has quadrupled in size, the axis labels, axis numbers, legend, etc. will stay the same size. This means everything will look really small when you increase the size of the graph, and really compressed when you decrease the size of the graph.

To deal with this, this package includes the function `ggsave_ifs()`, which is a wrapper for `ggsave()` that has a default height of 100mm, a default width of 150mm and saves at 320dpi.

Using the default options would be as so.

```{r example7, eval=FALSE}
ggsave_ifs(filename = "example1.png", plot = gg_object, path = "C:/this_directory")
```

If you want to manually adjust heights and widths (for example, if you have a graph with a lot going on), you can use:

```{r , eval=FALSE}
ggsave_ifs(filename = "example1.png", plot = gg_object, path = "C:/this_directory", height = 200, width = 300)
```

See the help file for more options for ggsave_ifs.
